TrueSkill Through Time: reliable initial skill estimates and historical comparability with \proglang{Julia}, \proglang{Python} and \proglang{R}

### 2022-02-02 doc/article.tex
doc.1.2.0: new version

\documentclass[article]{jss}
\usepackage[utf8]{inputenc}
% \usepackage{draftwatermark}
% \SetWatermarkText{Draft}
% \SetWatermarkScale{2}
\newif\ifen
\newif\ifes
\newcommand{\en}[1]{\ifen#1\fi}
\newcommand{\es}[1]{\ifes#1\fi}
\entrue
\usepackage[utf8]{inputenc}
\usepackage{paracol}
\usepackage{multirow}
\usepackage{booktabs}
%\newcommand{\jan}{%
%  \en{January}%
%  \es{Enero}
%}

@@

%% - \title{} in title case
%% - \Plaintitle{} without LaTeX markup (if any)
%% - \Shorttitle{} with LaTeX markup (if any), used as running title
\title{
%TrueSkill Through Time: \\ the \proglang{Julia}, \proglang{Python} and \proglang{R} packages
TrueSkill Through Time: reliable initial skill estimates and historical comparability in \proglang{Julia}, \proglang{Python} and \proglang{R}
}
\Plaintitle{TrueSkill Through Time: enabling reliable initial skill estimates and historical comparability}
\Shorttitle{TrueSkill Through Time}

@@

%% - \Abstract{} almost as usual
\Abstract{
\en{Humans develop complex skills through time. }%
\es{Los humanos desarrollamos habilidades complejas a través del tiempo. }%
%
\en{Knowing how individual abilities change is essential in a wide range of activities. }%
\es{Saber cómo cambian las capacidades individuales es esencial en varias actividades. }%
%
\en{Most estimators implemented for the video game industry cannot obtain reliable initial estimates or guarantee comparability between distant estimates because they do not properly model the temporal aspect of the learning process. }%
\es{La mayoría de los estimadores implementados para la industria del video juego no pueden obtener estimaciones iniciales fiables ni garantizar la comparabilidad entre estimaciones distantes debido a que no modelan adecuadamente el aspecto temporal del proceso de aprendizaje. }%
%
\en{Here we provide the first implementations of TrueSkill Through Time for \proglang{Julia}, \proglang{Python} and \proglang{R}, a model that solves all these problems by propagating historical information correctly, outperforming even more complex approaches. }%
\es{Aquí ofrecemos las primeras implementaciones de TrueSkill Through Time para \proglang{Julia}, \proglang{Python} y \proglang{R}, un modelo que resuelve todos estos problemas propagando correctamente la información histórica, superando incluso a enfoques similares más complejos. }%
%
\en{In such a sensitive problem as the estimation of people's ability, it is imperative to offer an explanation that is both intuitive and complete. }%
\es{En un problema tan sensible para las personas, como es la estimación de su habilidad es imperativo ofrecer una explicación que sea al mismo tiempo intuitiva y completa. }%
%
\en{In this paper we offer a technical report accessible to the general public without omitting any theoretical and technical details, material that is not available in the scientific literature. }%
\es{En este trabajo ofrecemos una reporte técnico accesible al público en general sin omitir ningún detalle téoricos y técnicos, material que no está disponible hasta ahora en la literatura científica. }%
%
\en{This algorithm requires only a few iterations to converge, allowing millions of observations to be analyzed using any low-end computer. }%
\es{Este algoritmo requiere sólo unas pocas iteraciones para converger, permitiendo analizar millones de observaciones usando cualquier ordenador de gama baja. }%
%
\en{After illustrating its basic mode of use, we show how to estimate the learning curves of historical players of the Association of Tennis Professionals (ATP). }%
\es{Luego de ilustrar su modo de uso básico, mostramos cómo estimar la curvas de aprendizaje de los jugadores históricas de la Asociación de Tenistas Profesionales (ATP). }%
%
\en{A complete scientific documentation allows scientists to make sense of all epistemological and technical aspects. }%
\es{Una documentación científica completa permite a los científicos dar sentido a todos los aspectos epistemológicos y técnicos. }%

} 
\Keywords{%
\en{Learning, skill, Bayesian inference, gaming, education, sports, \proglang{Julia}, \proglang{Python},  \proglang{R}}%
\es{Aprendizaje, habilidad, bayes, videojuego, educación, deporte, \proglang{Julia}, \proglang{Python},  \proglang{R}}%
}
\Plainkeywords{%
\en{Learning, skill, bayesian inference, gaming, education, sports, Julia, Python, R}%
\es{Aprendizaje, habilidad, bayes, videojuego, educación, deporte, Julia, Python, R}%
}

@@

    \edge {d} {r};
    \edge {p1,p2} {d};
    \edge {s1} {p1};
    \edge {s2} {p2};
    %\node[invisible, right=of p2, xshift=4.35cm] (s-dist) {};
}
     \caption{
     \en{Generative model in which skills cause the observable results mediated by the difference of hidden performances, $d =p_i - p_j$, both random variables around their unknown true skill, $p \sim \N(s,\beta^2)$. }%
    \es{Modelo generativo en el que las habilidades causan los resultados observables a trav\'es de la diferencia de rendimientos ocultos, $d=p_i-p_j$, ambas variables aleatorias centradas en la verdadera habilidad, $p \sim \N(s,\beta^2)$. }%
    %
    \en{The one with the highest performance wins, $r = (d > 0)$. }%
    \es{Quien haya obtenido mayor rendimiento gana, $r = (d > 0)$. }%
    %
    \en{Observable variables are painted gray, hidden in white, and constants are shown as black dots. }%
    \es{Las variables observables se pintan de gris, las ocultas en blanco, y las constantes se muestran como puntos negros. }%
    }
    \label{fig:generative_model}
\end{figure}
%
\en{Figure~\ref{fig:generative_model} provides a causal model in which skills generates the observable result. }%
\es{La figura~\ref{fig:generative_model} ofrece un modelo causal en el que las habilidades generan el resultado observable. }%
%
\en{The agents exhibit different performances at each event, varying around their true skill, $\N(p\,|\,s,\beta^2)$. }%
\es{Los agentes exhiben distintos desempe\~nos en cada evento, que var\'ian alrededor de su verdadera habilidad, $\N(p\,|\,s,\beta^2)$. }%
%
\en{The model assumes that the agent with the highest performance wins, $r = (p_i > p_j)$. In other words, whoever obtains a difference of performance greater than 0 wins, $r = (p_i - p_j > 0)$. }%
\es{El modelo supone que gana el agente con mayor rendimiento, $r = (p_i > p_j)$. En otras palabras, gana quien obtenga una diferencia de desempeño mayor a 0, $r = (p_i - p_j > 0)$. }%
%
\en{The parameter $\beta^2$ is the same for all the agents and acts as the scale of the estimates: skill estimates which differ by $\beta$ are associated with the same probability of winning independently of the absolute value of the estimates. }%
\es{El parámetro $\beta^2$, al ser el mismo para todos los agentes, act\'ua como la escala de las estimaciones: habilidades separada por un $\beta$ mantienen siempre la misma probabilidad de ganar independiente del valor absoluto de las estimaciones. }%

@@

\en{In table~\ref{Tab:Models} we compare our implementation of the TTT algorithm against Elo, TrueSkill, Constant and KickScore. The first two are versions of the ``filtering'' approaches and the second two are versions of the ``smoothing'' approach. }%
\es{En la tabla~\ref{Tab:Models} comparamos nuestra implementación del algoritmo TTT contra Elo, TrueSkill, Constant y  KickScore. Las primeras dos son versiones de los enfoques de ``filtering'' y las segundas dos son versiones de ``smoothing''. }%
\en{To evaluate the models we must calculate their probability given the data, }%
\es{Para evaluar los modelos debemos calcular su probabilidad dados los datos, }%
 P(\text{Model\es{o}}|\text{Dat\en{a}\es{os}}) = \frac{P(\text{Dat\en{a}\es{os}}|\text{Model\es{o}})P(\text{Model\es{o}})}{P(\text{Dat\en{a}\es{os}})}
\en{When we do not have access to all models, we cannot compute the probability $P(\text{Model}|\text{Data})$, but we can compare models. }%
\es{Cuando no tenemos acceso a todos los modelos, no podremos calcular la probabilidad $P(\text{Modelo}|\text{Datos})$, pero sí podremos comparar modelos. }%
\begin{equation}\label{eq:bayes_factor}
 \frac{P(\text{Model\es{o}}_i|\text{Dat\en{a}\es{os}})}{P(\text{Model\es{o}}_j|\text{Dat\en{a}\es{os}})} = \frac{P(\text{Dat\en{a}\es{os}}|\text{Model\es{o}}_i)P(\text{Model\es{o}}_i)}{P(\text{Dat\en{a}\es{os}}|\text{Model\es{o}}_j)P(\text{Model\es{o}}_j)} \overset{\text{\en{if}\es{si} } *}{=} \frac{P(\text{Dat\en{a}\es{os}}|\text{Model\es{o}}_i)}{P(\text{Dat\en{a}\es{os}}|\text{Model\es{o}}_j)} 
\en{This expression is known as the \emph{Bayes factor}, which is usually expressed on a logarithmic scale to report the difference in orders of magnitude. }%
\es{Esta expresión se conoce como el \emph{Bayes factor}, el cual se suele expresar en escala logaritmica para reportar la diferencia en órdenes de magnitud. }%
\en{In the special case where we have no prior preference over any model ($\overset{\text{\en{if}\es{si} } *}{=}$), we need only to compare the prior predictions of the models. }%
\es{En el caso especial en el que no tenemos preferencia a priori sobre ningún modelo ($\overset{\text{\en{if}\es{si} } *}{=}$), sólo debemos comparar las predicciones a priori de los modelos. }%
P(\text{Dat\en{a}\es{os}}|\text{Model\es{o}}) = P(d_1|\text{Model\es{o}})P(d_2|d_1,\text{Model\es{o}}) \dots P(d_n|d_1,\dots, d_{n-1},\text{Model\es{o}})
\en{where each element of the product is a prediction of the next data point based on the previous data set and the model. }%
\es{donde cada elementos de la productoria es una predicción del siguiente dato en base a los datos previos y el modelo. }%
\en{In this work, we report the geometric mean because it can be interpreted as the ``characteristic'' prediction of the model, }%
\es{En este trabajo también vamos a reportare la media geométrica debido a que puede interpretarse como la predicción ``característica'' del modelo, }%
\begin{equation}\label{eq:gmean}
P(\text{Dat\en{a}\es{os}}|\text{Model\es{o}}) = \text{geometric mean}(P(\text{Dat\en{a}\es{os}}|\text{Model\es{o}}))^{|\text{Dat\en{a}\es{os}}|}
\en{where $|\text{Data}|$ represents the size of the data set. }%
\es{donde $|\text{Datos}|$ representa el tamaño del conjunto de datos. }%
\en{To make the comparison we will follow the methodology employed by \cite{Maystre2019} over the four databases on which they report their prior predictions: ATP (tennis), NBA (basketball), FIFA (football), and FIDE (chess). }%

@@

\en{This methodology consists of splitting the data in two subsets: the initial \SI{70}{\percent} used to train the hyperparameters and the last \SI{30}{\percent} to evaluate the models. }%

@@

%ttt: 0.576047 %ks: 0.5758245946309177

@@

%ttt: 0.5317903539908228 %ks: 0.532802033269822 %ks_integrado: 0.5327914819          

@@

 %ttt: 0.358379593836443  %ks: 0.358397563993023 %ks_integrado: 0.3583680733

@@

 %ttt: 0.39627670749811733 ks: 0.39629097832609717 ks_integrado: 0.39619785744480546

@@

  \en{Comparisons between TTT, KickScore, TrueSkill, Elo and Constant models in four databases. }%

@@

  \en{The $\text{log}_2$BF columns report the Bayes Factor (Eq.~\ref{eq:bayes_factor}) between TTT and the other models in logarithmic scale. }%

@@

  \en{For reference, LOOCV column reports the geometric mean of the predictions using all historical data, $p(d_i| d_1, \dots, d_{i-1}, d_{i+1}, \dots, d_n , M)$. }%

@@

  \en{The geometric mean of the models indicated with the symbol $\dagger$ were computed by \cite{Maystre2019}. }%
  \es{La media geométrica de los modelos señalados con el símbolo $\dagger$ fueron computadas por \cite{Maystre2019}. }%

@@

\en{To compute the predictions~\cite{Maystre2019} use a HPC cluster to create, in parallel, a different model per day, each requiring about $100$ iterations to reach convergence. }%
\es{Para computar las predicciones ellos utilizan un cluster para crear en paralelo un modelo diferente por cada día $t$, que requieren cada uno cerca de 100 iteraciones para alcanzar convergencia. }%
\en{Instead, it was sufficient for us to use a desktop computer to create a single model adding data day by day and performing a single iteration at each step. }%
\es{En cambio, para nosotros fue suficiente usar una computadora de escritorio para crear un único modelo al que le agregamos las datos día a día, realizando una única iteración en cada paso. }%
\en{The model selection was made by optimizing the hyperparameters, but a full Bayesian model selection should compute predictions by integrating the entire hyperparameter space.  }%
\es{La selección de modelo la hemos realizado optimizando los hiperparámetros, pero una selección de modelo bayesiana completa debería computar las predicciones integrando todo el espacio de hiperparámetros. }%
\en{This procedure usually penalizes complex models when the search in the hyperparameter space is unnecessary. }%
\es{Este procedimiento suele penalizar a los modelos complejos cuando la búsqueda en el espacio de hiperparámetros es innecesaria. }%
\en{With no more than 3 intuitive hyperparameters, the TTT model achieves similar or even better results than KickScore more efficiently, because the latter must optimize on a much larger hyperparameter space, which includes the definition of a Kernel (a statistical trick not always intuitive\footnote{\href{https://stats.stackexchange.com/questions/152897/how-to-intuitively-explain-what-a-kernel-is}{Stackexchange: ``How to intuitively explain what a kernel is.'' (question 152897)}}) and then the hyperparameters specific to that Kernel. }%
\es{Con no más de 3 hiperparámetros intuitivos, el modelo TTT logra resultados similares y hasta mejores que KickScore de forma más eficiente, debido a que el último debe optimizar en un espacio de parámetros mucho más grande, que incluye la definción de un Kernel (un truco estadísitico no siempre intuitivo\footnote{\href{https://stats.stackexchange.com/questions/152897/how-to-intuitively-explain-what-a-kernel-is}{Stackexchange: ``How to intuitively explain what a kernel is.'' (question 152897)}}) y luego los hiperparámetros específicos a ese Kernel. }%

\en{Often the objective is to estimate learning curves over time and not the online predictions that we have computed to compare models. }%
\es{Muchas veces el objetivo es estimar las curvas de aprendizaje en el tiempo y no las predicciones en línea que hemos computado para comparar modelos. }%
%
\en{In these cases we want to find the best estimates of skill, from the first to the last event of the learning curve. }%
\es{En estos casos queremos encontrar las mejores estimaciones de habilidad, desde el primer hasta el último evento de la curva de aprendizaje. }%
%
\en{Online predictions generate poor initial estimates because we force the model to perform inference using only past information. }%
\es{Las predicciones en línea generan malas estimaciones iniciales debido a que forzamos al modelo a realizar la inferencia usando solamente la información del pasado. }%
%
\en{However, nothing prevents probabilistic models from making inference using information from the future, something we intuitively do when we suspect that prominent athletes were also skilled some time before they became famous. }%
\es{Sin embargo, nada impide que los modelos probabilísiticos realicen inferencia usando información del futuro, algo que intuitivamente hacemos cuando sospechamos que los deportistas destacados también eran hábiles un tiempo antes de hacerse conocidos. }%
%
\en{The LOOCV column of Table~\ref{Tab:Models} reports the geometric mean of the predictions when the skill estimation is performed using all historical information except the event we are predicting, $p(d_i| d_1, \dots, d_{i-1}, d_{i+1}, \dots, d_n , M)$. }.
\es{La columna LOOCV de la tabla~\ref{Tab:Models} reporta la media geométrica de las predicciones cuando la estimación de la habilidades se realiza usando toda la información histórica salvo el evento que estamos prediciendo, $p(d_i| d_1, \dots, d_{i-1}, d_{i+1}, \dots, d_n , M)$. }%
%
\en{While we do not calculate the Bayes Factor, these estimates are hundreds of orders of magnitude better than estimates made using only past information. }%
\es{Si bien no calculamos el Bayes Factor, estas estimaciones son centenas de órdenes de magnitud mejores que las estimaciones hechas usando sólo la información del pasado. }%

\en{In such a sensitive problem as the estimation of people's ability, it is imperative to offer an explanation that is both intuitive and complete. }%
\es{En un problema tan sensible para las personas, como es la estimación de su habilidad es imperativo ofrecer una explicación que sea al mismo tiempo intuitiva y completa. }%
%
\en{There is a very large population, both in the video game industry and in the educational systems, that needs to know how the algorithm that measures their skills works. }%
\es{Hay una población muy grande, tanto en la industria del video juego como en los sistemas educativos, que necesita conocer cómo funciona el algoritmo que mide su habilidad. }%
%
\en{This paper aims to provide a technical report, with all the theoretical and technical details, in the most accessible way possible for the general public. }%
\es{El objetivo de este trabajo es ofrecer una reporte técnico con todos los detalles téoricos y técnicos de la forma más accesible posible al público en general. }%

@@

\en{However, there are also temporal and spatial decouplings that make it impossible to compare estimates: although the relative differences between current estimates within well-connected communities are correct, estimates separated in time and between poorly connected communities are often incorrect. }%
\es{Pero tambi\'en se observan desacoplamientos temporales y espaciales que impiden comparar estimaciones: aunque la diferencias relativas entre estimaciones contempor\'aneas al interior de comunidades bien conectadas sean correctas, las estimaciones separadas en el tiempo y entre comunidades poco conectadas suelen ser incorrectas. }%

@@

\en{In a brief, the messages sent by the variable-type node $v$ are simply the product of the messages that $v$ receives from the rest of their neighbors $h \in n(v)$ except $f$. }%
\es{En pocas palabras, los mensajes que env\'ia una variables $v$ es simplemente la multiplicaci\'on de los mensajes que recibi\'o del resto de sus vecinos $h \in n(v)$ salvo $f$. }%

@@

\en{The messages sent by the function-type node $f$ to a neighboring variable-type node $v$ is simply the integration (or sum) over $\bm{h}$ of the product of itself and all the messages that $f$ receives from the rest of its neighbors $\bm{h}$ except $v$. }%

>>>>

\documentclass[article]{jss}
\usepackage[utf8]{inputenc}
% \usepackage{draftwatermark}
% \SetWatermarkText{Draft}
% \SetWatermarkScale{2}
\newif\ifen
\newif\ifes
\newcommand{\en}[1]{\ifen#1\fi}
\newcommand{\es}[1]{\ifes#1\fi}
\entrue
\usepackage[utf8]{inputenc}
\usepackage{paracol}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{subcaption}
%\newcommand{\jan}{%
%  \en{January}%
%  \es{Enero}
%}

@@

%% - \title{} in title case
%% - \Plaintitle{} without LaTeX markup (if any)
%% - \Shorttitle{} with LaTeX markup (if any), used as running title
\title{
%TrueSkill Through Time: \\ the \proglang{Julia}, \proglang{Python} and \proglang{R} packages
TrueSkill Through Time: reliable initial skill estimates and historical comparability with \proglang{Julia}, \proglang{Python} and \proglang{R}
}
\Plaintitle{TrueSkill Through Time: reliable initial skill estimates and historical comparability with Julia, Python and R}
\Shorttitle{TrueSkill Through Time}

@@

%% - \Abstract{} almost as usual
\Abstract{
\en{Humans develop complex skills through time. }%
\es{Los humanos desarrollamos habilidades complejas a través del tiempo. }%
\en{Knowing how individual abilities change is essential in a wide range of activities. }%
\es{Saber cómo cambian las capacidades individuales es esencial en varias actividades. }%
\en{The most commonly used skill estimators in the video game industry (such as Elo and TrueSkill) cannot obtain reliable initial estimates or guarantee comparability between estimates distant in time and space. }%because they do not propagate historical information properly through the system. }%
\es{Los estimadores de habilidad más utilizados en la industria del video juego (como Elo y TrueSkill) no pueden obtener estimaciones iniciales fiables ni garantizar la comparabilidad entre estimaciones distantes en el tiempo y el espacio. }%debido a que no propagan la información histórica apropiadamente a través del sistema. }%
\en{The model TrueSkill Through Time (TTT) propagates all historical information throughout the system, enabling reliable initial skill estimates and guaranteeing historical comparability. }%
\es{El modelo TrueSkill Through Time (TTT) propaga toda la información histórica a través del sistema, lo que permite realizar estimaciones fiables de la habilidad inicial y garantizar la comparabilidad histórica. }%
\en{The improvement is evident from the prior predictions, which are several orders of magnitude more accurate than those obtained with the aforementioned models. }%
\es{La mejora se hace evidente a partir de las predicciones a priori, que son varios órdenes de magnitud más precisas que las obtenidas con los modelos mencionados. }%
\en{With no more than three intuitive hyperparameters, the TTT model achieves similar and even better results than more complex models such as KickScore, in a more efficient way. }%
\es{Con no más de tres hiperparámetros intuitivos, el modelo TTT logra resultados similares y hasta mejores que modelos más complejos como KickScore, de forma más eficiente. }%
\en{Although the TTT model was published more than a decade ago, it was not available until now in the programming languages with the largest communities. }%
\es{Si bien el modelo TTT fue publicado hace más una década, no se encontraba disponible hasta ahora en los lenguajes de programación con mayores comunidades. }%
\en{Here we offer the first software for \proglang{Julia}, \proglang{Python} and \proglang{R}, accompanied by an intuitive explanation for the general public and an in-depth explanation to people with basic knowledge in probability. }%
\es{Aquí ofrecemos el primer software para \proglang{Julia}, \proglang{Python} y \proglang{R}, acompañado por una explicación intuitiva comprensible para el público general, y otra explicación en profundidad específica para las persona con conocimientos básicos en probabilidad. }%
%%
\en{After illustrating its basic mode of use, we show how to estimate the learning curves of historical players of the Association of Tennis Professionals (ATP). }%
\es{Luego de ilustrar su modo de uso básico, mostramos cómo estimar la curvas de aprendizaje de los jugadores históricas de la Asociación de Tenistas Profesionales (ATP). }%
\en{This algorithm requires few iterations to converge, allowing millions of observations to be analyzed using any low-end computer. }%
\es{Este algoritmo requiere pocas iteraciones para converger, permitiendo analizar millones de observaciones usando cualquier ordenador de gama baja. }%
\Keywords{%
\en{Learning, skill, Bayesian inference, gaming, education, sports, \proglang{Julia}, \proglang{Python},  \proglang{R}}%
\es{Aprendizaje, habilidad, bayes, videojuego, educación, deporte, \proglang{Julia}, \proglang{Python},  \proglang{R}}%
\Plainkeywords{%
\en{Learning, skill, bayesian inference, gaming, education, sports, Julia, Python, R}%
\es{Aprendizaje, habilidad, bayes, videojuego, educación, deporte, Julia, Python, R}%

@@

    \edge {d} {r};
    \edge {p1,p2} {d};
    \edge {s1} {p1};
    \edge {s2} {p2};

}
     \caption{
     \en{Generative model in which skills cause the observable results mediated by the difference of hidden performances, $d =p_i - p_j$, both random variables around their unknown true skill, $p \sim \N(s,\beta^2)$. }%
    \es{Modelo generativo en el que las habilidades causan los resultados observables a trav\'es de la diferencia de rendimientos ocultos, $d=p_i-p_j$, ambas variables aleatorias centradas en la verdadera habilidad, $p \sim \N(s,\beta^2)$. }%
    %
    \en{The one with the highest performance wins, $r = (d > 0)$. }%
    \es{Quien haya obtenido mayor rendimiento gana, $r = (d > 0)$. }%
    %
    \en{Observable variables are painted gray, hidden in white, and constants are shown as black dots. }%
    \es{Las variables observables se pintan de gris, las ocultas en blanco, y las constantes se muestran como puntos negros. }%
    }
    \label{fig:generative_model}
\end{figure}
%
\en{Figure~\ref{fig:generative_model} provides a causal model in which skills generates the observable result. }%
\es{La figura~\ref{fig:generative_model} ofrece un modelo causal en el que las habilidades generan el resultado observable. }%
%
\en{The agents exhibit different performances at each event, varying around their true skill, $\N(p\,|\,s,\beta^2)$. }%
\es{Los agentes exhiben distintos desempe\~nos en cada evento, que var\'ian alrededor de su verdadera habilidad, $\N(p\,|\,s,\beta^2)$. }%
%
\en{The model assumes that the agent with the highest performance wins, $r = (p_i > p_j)$. In other words, whoever obtains a difference of performance greater than 0 wins, $r = (p_i - p_j > 0)$. }%
\es{El modelo supone que gana el agente con mayor rendimiento, $r = (p_i > p_j)$. En otras palabras, gana quien obtenga una diferencia de desempeño mayor a 0, $r = (p_i - p_j > 0)$. }%
%
\en{The parameter $\beta^2$ is the same for all the agents and acts as the scale of the estimates: skill estimates which differ by $\beta$ are associated with the same probability of winning independently of the absolute value of the estimates. }%
\es{El parámetro $\beta^2$, al ser el mismo para todos los agentes, act\'ua como la escala de las estimaciones: habilidades separada por un $\beta$ mantienen siempre la misma probabilidad de ganar independiente del valor absoluto de las estimaciones. }%

@@

\en{In table~\ref{Tab:Models} we compare our implementation of the TTT algorithm against Elo, TrueSkill, Constant and KickScore. The first two are versions of the ``filtering'' approaches and the second two are versions of the ``smoothing'' approach. }%
\es{En la tabla~\ref{Tab:Models} comparamos nuestra implementación del algoritmo TTT contra Elo, TrueSkill, Constant y  KickScore. Las primeras dos son versiones de los enfoques de ``filtering'' y las segundas dos son versiones de ``smoothing''. }%
%
\en{To evaluate the models we must calculate their probability given the data, }%
\es{Para evaluar los modelos debemos calcular su probabilidad dados los datos, }%
%
\begin{equation}
 P(\text{Model\es{o}}|\text{Dat\en{a}\es{os}}) = \frac{P(\text{Dat\en{a}\es{os}}|\text{Model\es{o}})P(\text{Model\es{o}})}{P(\text{Dat\en{a}\es{os}})}
\end{equation}
%
\en{When we do not have access to all models, we cannot compute the probability $P(\text{Model}|\text{Data})$, but we can compare models. }%
\es{Cuando no tenemos acceso a todos los modelos, no podremos calcular la probabilidad $P(\text{Modelo}|\text{Datos})$, pero sí podremos comparar modelos. }%
%
\begin{equation}\label{eq:bayes_factor}
 \frac{P(\text{Model\es{o}}_i|\text{Dat\en{a}\es{os}})}{P(\text{Model\es{o}}_j|\text{Dat\en{a}\es{os}})} = \frac{P(\text{Dat\en{a}\es{os}}|\text{Model\es{o}}_i)P(\text{Model\es{o}}_i)}{P(\text{Dat\en{a}\es{os}}|\text{Model\es{o}}_j)P(\text{Model\es{o}}_j)} \overset{\text{\en{if}\es{si} } *}{=} \frac{P(\text{Dat\en{a}\es{os}}|\text{Model\es{o}}_i)}{P(\text{Dat\en{a}\es{os}}|\text{Model\es{o}}_j)} 
\end{equation}
%
\en{This expression is known as the \emph{Bayes factor}, which is usually expressed on a logarithmic scale to report the difference in orders of magnitude. }%
\es{Esta expresión se conoce como el \emph{Bayes factor}, el cual se suele expresar en escala logaritmica para reportar la diferencia en órdenes de magnitud. }%
%
\en{In the special case where we have no prior preference over any model ($\overset{\text{\en{if}\es{si} } *}{=}$), we need only to compare the prior predictions of the models. }%
\es{En el caso especial en el que no tenemos preferencia a priori sobre ningún modelo ($\overset{\text{\en{if}\es{si} } *}{=}$), sólo debemos comparar las predicciones a priori de los modelos. }%
%
\begin{equation}
P(\text{Dat\en{a}\es{os}}|\text{Model\es{o}}) = P(d_1|\text{Model\es{o}})P(d_2|d_1,\text{Model\es{o}}) \dots P(d_n|d_1,\dots, d_{n-1},\text{Model\es{o}})
\end{equation}
%
\en{where each element of the product is a prediction of the next data point based on the previous data set and the model. }%
\es{donde cada elementos de la productoria es una predicción del siguiente dato en base a los datos previos y el modelo. }%
%
\en{In this work, we report the geometric mean because it can be interpreted as the ``characteristic'' prediction of the model, }%
\es{En este trabajo también vamos a reportare la media geométrica debido a que puede interpretarse como la predicción ``característica'' del modelo, }%
%
\begin{equation}\label{eq:gmean}
P(\text{Dat\en{a}\es{os}}|\text{Model\es{o}}) = \text{geometric mean}(P(\text{Dat\en{a}\es{os}}|\text{Model\es{o}}))^{|\text{Dat\en{a}\es{os}}|}
\end{equation}
%
\en{where $|\text{Data}|$ represents the size of the data set. }%
\es{donde $|\text{Datos}|$ representa el tamaño del conjunto de datos. }%

\en{We follow the methodology employed by \cite{Maystre2019} to compare the different models using four databases in which they report the prior predictions: ATP (tennis), NBA (basketball), FIFA (football), and FIDE (chess). }%

@@

\en{This methodology consists of splitting the data into two subsets: the initial \SI{70}{\percent} used to train the hyperparameters and the last \SI{30}{\percent} to evaluate the models. }%

@@

@@

@@

@@

@@

  \en{Comparisons between different models in four databases. }%

@@

  \en{The $\log_2$BF columns report the Bayes Factor (Eq.~\ref{eq:bayes_factor}) between TTT and the other models in the logarithmic scale. }%

@@

  \en{For reference, the LOOCV column reports the geometric mean of the predictions using all historical data, $p(d_i| d_1, \dots, d_{i-1}, d_{i+1}, \dots, d_n , M)$. }%

@@

  \en{The geometric mean of the models marked with the symbol $\dagger$ was computed by \cite{Maystre2019}. }%
  \es{La media geométrica de los modelos señalados con el símbolo $\dagger$ fue obtenida por \cite{Maystre2019}. }%

@@

\en{To compute the predictions~\cite{Maystre2019} use a HPC cluster to create, in parallel, a different model per day, each requiring about $100$ iterations to reach convergence. }%
\es{Para computar las predicciones ellos utilizan un cluster para crear en paralelo un modelo diferente por cada día $t$, que requieren cada uno cerca de 100 iteraciones para alcanzar convergencia. }%
%
\en{Instead, it was sufficient for us to use a desktop computer to create a single model adding data day by day and performing a single iteration at each step. }%
\es{En cambio, para nosotros fue suficiente usar una computadora de escritorio para crear un único modelo al que le agregamos las datos día a día, realizando una única iteración en cada paso. }%
%
\en{The model selection was made by optimizing the hyperparameters, but a full Bayesian model selection should compute predictions by integrating the entire hyperparameter space.  }%
\es{La selección de modelo la hemos realizado optimizando los hiperparámetros, pero una selección de modelo bayesiana completa debería computar las predicciones integrando todo el espacio de hiperparámetros. }%
%
\en{This procedure usually penalizes complex models when the search in the hyperparameter space is unnecessary. }%
\es{Este procedimiento suele penalizar a los modelos complejos cuando la búsqueda en el espacio de hiperparámetros es innecesaria. }%
%
\en{With no more than three intuitive hyperparameters, the TTT model achieves similar or even better results than KickScore more efficiently, because the latter must optimize on a much larger hyperparameter space, which includes the definition of a kernel and then the hyperparameters specific to that kernel. }%
\es{Con no más de tres hiperparámetros intuitivos, el modelo TTT logra resultados similares y hasta mejores que KickScore de forma más eficiente, debido a que el último debe optimizar en un espacio de parámetros mucho más grande, que incluye la definción de un \emph{kernel} y luego los hiperparámetros específicos a ese \emph{kernel}. }%

\en{Often the objective is to estimate learning curves over time and not the online predictions that we have computed to compare models. }%
\es{Muchas veces el objetivo es estimar las curvas de aprendizaje en el tiempo y no las predicciones en línea que hemos computado para comparar modelos. }%
%
\en{In these cases we want to find the best estimates of skill, from the first to the last event of the learning curve. }%
\es{En estos casos queremos encontrar las mejores estimaciones de habilidad, desde el primer hasta el último evento de la curva de aprendizaje. }%
%
\en{Online predictions generate poor initial estimates because we force the model to perform inference using only past information. }%
\es{Las predicciones en línea generan malas estimaciones iniciales debido a que forzamos al modelo a realizar la inferencia usando solamente la información del pasado. }%
%
\en{However, nothing prevents probabilistic models from making inference using information from the future, something we intuitively do when we suspect that prominent athletes were also skilled some time before they became famous. }%
\es{Sin embargo, nada impide que los modelos probabilísiticos realicen inferencia usando información del futuro, algo que intuitivamente hacemos cuando sospechamos que los deportistas destacados también eran hábiles un tiempo antes de hacerse conocidos. }%
%
\en{The LOOCV column of Table~\ref{Tab:Models} reports the geometric mean of the predictions when the skill estimation is performed using all historical information except the event we are predicting, $p(d_i| d_1, \dots, d_{i-1}, d_{i+1}, \dots, d_n , M)$. }%
\es{La columna LOOCV de la tabla~\ref{Tab:Models} reporta la media geométrica de las predicciones cuando la estimación de la habilidades se realiza usando toda la información histórica salvo el evento que estamos prediciendo, $p(d_i| d_1, \dots, d_{i-1}, d_{i+1}, \dots, d_n , M)$. }%
%
\en{While we do not calculate the Bayes Factor, these estimates are hundreds of orders of magnitude better than estimates made using only past information. }%
\es{Si bien no calculamos el Bayes Factor, estas estimaciones son centenas de órdenes de magnitud mejores que las estimaciones hechas usando sólo la información del pasado. }%

\en{In such a sensitive problem as the estimation of people's ability, it is imperative to offer an explanation that is both intuitive and complete. }%
\es{En un problema tan sensible para las personas, como es la estimación de su habilidad es imperativo ofrecer una explicación que sea al mismo tiempo intuitiva y completa. }%
%
\en{Large communities like the video game industry and educational systems, need to be sure that the algorithm estimating their skills actually works. }%
\es{Hay una población muy grande, tanto en la industria del video juego como en los sistemas educativos, que necesita conocer cómo funciona el algoritmo que mide su habilidad. }%
%
\en{This paper aims to present a working tool jointly with all the theoretical and technical details to show the foundations of the method. }%
\es{El objetivo de este trabajo es ofrecer una reporte técnico con todos los detalles téoricos y técnicos de la forma más accesible posible al público en general. }%

@@

\en{But temporal and spatial decouplings may also occur, preventing the comparison between distant estimates. }%
\es{Pero también pueden producirse desacoplamientos temporales y espaciales que impidan comparar estimaciones distantes. }%
\en{Although the relative differences between current estimates within well-connected communities are correct, estimates separated in time and between poorly connected communities may be incorrect. }%
\es{Aunque la diferencias relativas entre estimaciones contempor\'aneas al interior de comunidades bien conectadas sean correctas, las estimaciones separadas en el tiempo y entre comunidades poco conectadas pueden ser incorrectas. }%

@@

\en{In brief, the messages sent by the variable-type node $v$ are simply the product of the messages that $v$ receives from the rest of their neighbors $h \in n(v)$ except $f$. }%
\es{En pocas palabras, los mensajes que env\'ia una variable $v$ es simplemente la multiplicaci\'on de los mensajes que recibi\'o del resto de sus vecinos $h \in n(v)$ salvo $f$. }%

@@

\en{In brief, the messages sent by a function $f$ is also the product of the messages it receives from the rest of its neighbors $h \in n(f)$ except $v$, but this time it is also multiplied by itself $f(\cdot)$ and integrated (or summed) over $\bm{h}$. }%
\es{En pocas palabras, los mensajes que envía una funci\'on $f$ también es la multiplicación de los mensajes que recibi\'o del resto de sus vecinos $h \in n(f)$ salvo $v$, por sí misma $f(\cdot)$, integrando (o sumando) sobre $\bm{h}$. }%
